Closure-Free Functional Programming in a Two-Level Type Theory

### 2024-06-12 newpaper/paper.tex
wip

\documentclass[acmsmall,screen,review,anonymous]{acmart}
%% \documentclass[nonacm,acmsmall,review]{acmart}
%% \raggedbottom

@@

%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{András Kovács}
\email{andrask@chalmers.se}
\orcid{0000-0002-6375-9781}
\affiliation{%
  \institution{University of Gothenburg}
  \country{Sweden}
  \city{Gothenburg}
}

@@

Why not just make the object language less liberal, e.g.\ by disallowing
$\lambda$ under $\mbf{case}$ or $\mbf{let}$, thereby making call saturation
easier or more obvious? There is a trade-off between making the object language
more restricted, and thus easier to compile, and making \emph{metaprogramming}
more convenient. We will see that the ability to insert $\mbf{let}$-s without
restriction is very convenient in code generation, and likewise the ability to
have arbitrary object expressions in $\mbf{case}$ bodies. In this paper we go
with the most liberal object syntax, at the cost of needing more downstream
processing.

On the other side of the spectrum, one might imagine generating object code in
low-level A-normal form. This is more laborious, but it could be also
interesting, because it forces us to invent abstractions for manipulating ANF in
the metalanguage.

>>>>

%% \documentclass[acmsmall,screen,review,anonymous]{acmart}
%% \documentclass[nonacm,acmsmall]{acmart}
\documentclass[acmsmall,screen]{acmart}
%% \raggedbottom

@@

%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{First Last}
\affiliation{%
  \institution{Institution}
  \country{Country}
  \city{City}
}

%% \author{András Kovács}
%% \email{andrask@chalmers.se}
%% \orcid{0000-0002-6375-9781}
%% \affiliation{%
%%   \institution{University of Gothenburg}
%%   \country{Sweden}
%%   \city{Gothenburg}
%% }

@@

Why not just make the object language less liberal, e.g.\ by disallowing
$\lambda$ under $\mbf{case}$ or $\mbf{let}$, thereby making call saturation
easier or more obvious? There is a trade-off between making the object language
more restricted, and thus easier to compile, and making \emph{metaprogramming}
more convenient. We will see that the ability to insert $\mbf{let}$-s without
restriction is very convenient in code generation, and likewise the ability to
have arbitrary object expressions in $\mbf{case}$ bodies. In this paper we go
with the most liberal object syntax, at the cost of needing more downstream
processing. On the other side of the spectrum, one might imagine generating
object code in low-level A-normal form. This is more laborious, but it could be
also interesting, because it would force us to invent abstractions for
manipulating ANF in the metalanguage. We leave such lower-level object languages
for future investigation.
